### 2022-06-08 manuscript.tex
Changing default CSL to Nature EcoEvol CSL [ci skip]
This build is based on
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/eafc396e1f32d378749c27716bfc058f4802928f.

This commit was created by the following CI build and job:
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/eafc396e1f32d378749c27716bfc058f4802928f/checks
https://github.com/SORTEE-Github-Hackathon/manuscript/actions/runs/2463521591

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/402617f9e4b0caf2a3e126b79bf17059d3cfef8c/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/402617f9e4b0caf2a3e126b79bf17059d3cfef8c}{SORTEE-Github-Hackathon/manuscript@402617f}
on June 8, 2022.

@@

Most scientists, including ecologists and evolutionary biologists, are increasingly dependent on computational tools in their research {[}\protect\hyperlink{ref-fJWFe93e}{1}{]}.
Researchers write and use code as part of their scientific workflow to perform a wide-variety of tasks ranging from data management, data analysis, and study replication, to the application and the development of tools for hypothesis testing.
To maintain code for scientific collaboration requires an efficient and well-documented work-flow {[}\protect\hyperlink{ref-1Kqna6l2}{2}{]}.
To facilitate this process, scientists have been increasingly adopting tools from information and systems technology, such as cloud-based services for documentation and version control (\emph{e.g.}, from the Google Suite, with Docs, Sheets, and Drive; the Microsoft Suite, with Word, Excel, and OneDrive; and GitHub; {[}\protect\hyperlink{ref-10ghgV3S8}{3}{]}).
However, most researchers lack exposure to adequate software development practices and are required to dedicate valuable time and effort to self-teach the use of research-facilitating tools, and thus may be limited in their ability to adhere to adequate standards of scientific code quality and maintenance {[}\protect\hyperlink{ref-fJWFe93e}{1},\protect\hyperlink{ref-10SpoByIw}{4},\protect\hyperlink{ref-O6UbstGG}{5}{]}.
Here, we review and discuss one of the most used web-based platforms for computational version control and collaboration, GitHub, and provide researchers in ecology and evolutionary biology (EEB) with practical workflows aimed at facilitating their scientific code and management process.

With over 73 million registered users, as of 2022, GitHub is the most widely-used web platform for collaborating on computer code {[}\protect\hyperlink{ref-nwCtHDCn}{6}{]}.
GitHub provides a simplified but powerful web interface that allows users to participate in projects, contribute code, report and discuss software bugs, discover existing code and data, and publish new code.
Through version control, users have a detailed, chronological record of the files and directories stored in their repositories (see \protect\hyperlink{definitions}{Box 1}; {[}\protect\hyperlink{ref-RVetqmsg}{7}{]}).
This workflow provides a strong and clear advantage over receiving, processing and sending files back-and-forth (e.g.~via email), a process that can easily become challenging and time-consuming in projects extending in time and in the number of collaborators {[}\protect\hyperlink{ref-4ny1onB0}{8}{]}.
Through the combination of version control management and the network- and collaboration-based features, GitHub can broadly facilitate openly available source code alongside concomitant collaborative development {[}\protect\hyperlink{ref-kEX5dgzK}{9}{]}.

Git is the version control software system that enables all the collaborative tools available on GitHub.
Although the understanding of basic concepts of Git (such as commit, push, pull, checkout; see \protect\hyperlink{definitions}{Box 1}) is necessary, the GitHub web-based platform and its integrated development environments (such as the GitHub Desktop) allow users to perform most repository and data management operations without opening Git command-line sessions.
Still, GitHub users are encouraged to explore and improve their proficiency in Git when feeling comfortable, so they can feel confident in performing more flexible, complex operations (such as integrating changes from one branch into another branch with \texttt{git\ merge} and \texttt{git\ rebase}). Examples of extensive explanations on Git can be found in journal articles {[}\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-PlcxShQU}{10}{]}, video tutorials {[}\protect\hyperlink{ref-NIS0JOW0}{11}{]}, and books {[}\protect\hyperlink{ref-13jOlVcpp}{12}{]}.

The expansive GitHub user-community and the numerous resources on how to use GitHub have boosted its growing popularity {[}\protect\hyperlink{ref-RVetqmsg}{7},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-13jOlVcpp}{12}/,\protect\hyperlink{ref-u5aEVE4B}{13}{]}.
Nevertheless, although multiple articles have encouraged researchers in EEB to adopt GitHub as part of their research process {[}\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-3DKwn1sY}{14}{]}, its use is still not widespread.
First-time users without formal training in information technology may face steep learning curves because GitHub and its features have been centered on collaboration for software development in information systems {[}\protect\hyperlink{ref-139b0pSGc}{15}{]}.
Moreover, domain-specific perspectives and resources providing tractable examples and practical guidance for researchers in EEB on GitHub are scarce (but see https://ourcodingclub.github.io; https://www.openscapes.org).
A common adoption of GitHub for collaborating on a variety of research tasks can ultimately enable EEB researchers to spend less time on creating novel processes for collaboration and more time on their research {[}\protect\hyperlink{ref-ydrk01SR}{16}{]}.
More importantly, increasing the availability of data and code management standards -- of which GitHub is one increasingly important component -- make research more reproducible and collaborative {[}\protect\hyperlink{ref-13QX8XU3J}{17}{]}.

@@

In EEB, a research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Having multiple copies of these files is important to ensure that research can continue should one copy of a file be accidentally modified or deleted.
Many researchers begin using GitHub as a means to backup their research compendium {[}\protect\hyperlink{ref-MwwMapRG}{18}{]} to a remote server (pull-commit-push, see \protect\hyperlink{definitions}{Box 1}).
The GitHub repository serves as a centralized backup and also allows users to synchronize files and work on any device with internet access.
Because GitHub is built on git, version control features are also available, allowing users to re-visit or even restore previous versions of the repository.
Storing a research compendium as a GitHub repository has additional benefits such as facilitating collaboration, integration with data and code archiving services, and contribution to open science, all discussed in sections below.

Unfortunately, GitHub repositories have file size limits, so they may not be appropriate for synchronizing large data files.
Commits (see \protect\hyperlink{definitions}{Box 1}{]}) greater than 50 MB receive a warning and commits larger than 100 MB are blocked ({[}\protect\hyperlink{ref-1Co6ZZjF1}{19}{]}).
Therefore, it is recommended that large datasets are stored only locally or using other cloud services.

@@

Lab notebooks are the virtual or physical notebooks used to help researchers at any career stage keep track of their research methods or laboratory policies {[}\protect\hyperlink{ref-wwHxTOtm}{20}{]}.
Digital lab notebooks stored in the cloud provide clear benefits given the ease with which documents can be shared with new employees and updated as policy changes or experimental methods are modified {[}\protect\hyperlink{ref-10V7x4H4l}{21}{]}.
Increasingly, researchers are leveraging GitHub's underlying version control system to maintain and share digital lab notebooks {[}\protect\hyperlink{ref-10ghgV3S8}{3}{]}.
At least for aspects of a research project that involve writing code, a GitHub repository is a form of a lab notebook; when changes are made to files in a version controlled repository, the author of those changes makes a commit (\protect\hyperlink{definitions}{Box 1}) accompanied by a message describing the changes and the reason for them. Later, the entire history of commits and their messages are viewable and can be audited similar to a physical lab notebook {[}\protect\hyperlink{ref-4ny1onB0}{8}{]}.
GitHub issues (\protect\hyperlink{definitions}{Box 1}) can be used to prioritize lab objectives and goals, as well as track any status updates.
Some EEB labs have even turned their lab notebooks into shareable websites (\url{https://scheuerell-lab.github.io/lab-book}; \url{https://github.com/HuckleyLab/how_we_work}) as a centralized location for all lab resources.

@@

Modern EEB research is highly collaborative, bringing together multidisciplinary teams from various institutions.
On GitHub, collaborators can share feedback, brainstorm ideas, and troubleshoot problems (Figure \ref{fig:github-diagram}).
GitHub has an ``Issues'' feature that allows for discrete tasks and sub-tasks to be identified, assigned to team members, and categorized with custom labels.
The new GitHub ``Discussion'' feature serves as a message board for conversation.
Scripts, commit messages, and pull requests can be linked directly to issues and discussions, providing a clear record of project workflow.
The use of GitHub for all project-related conversation and planning, rather than e-mail or messaging tools, makes it easier to keep track of progress throughout the lifespan of a project.
Unlike emails and messages which can get lost as more new tasks arise, GitHub issues are intentionally closed by repository administrators hiding the issue from view (closed issues remain accessible but not immediately visible).
Fortunately, it is not essential for all team members to have proficiency in git or programming, as users can interact with Issues and Discussions via web browser or e-mail (e-mail responses still get tracked as comments on the focal GitHub issue).
For larger projects with many team members and tens or hundreds of GitHub issues to sort through, project management software like ZenHub, can help prioritize issues and pull requests.
ZenHub's web interface includes a GitHub Issue visualizer where users can organize issues into high priority or backlogged tasks and link issues together when they are related to a shared project goal or milestone.
GitHub is currently beta testing a similar project management feature called GitHub Projects {[}\protect\hyperlink{ref-RhBKe0MG}{22}{]}. GitHub can also be integrated with other project management software such as Slack (https://slack.github.com/) so that teams are notified through a Slack group chat when updates are made to a repository.

@@

Making presentations can be done through most major high-level programming languages such as R, with \texttt{RMarkdown} {[}\protect\hyperlink{ref-QqMezOMg}{23}{]}, Python, with \texttt{python-ppt} (\url{https://python-pptx.readthedocs.io}), and Julia, with \texttt{Remark.jl} (\url{https://juliapackages.com/p/remark}).
Since all these programs work via code bases, they can be version-controlled through git and GitHub.
Once content is made, hosting a course website can be done through GitHub Pages, and there are lots of templates available (e.g., see \url{https://github.com/topics/course-website}).
This way, the course content can be available to enrolled students, as well as a global pool of learners and teachers interested in the course material.
Content can then be delivered via the course website, and/or a GitHub organization with, for example, template repositories for assignments.
Student submissions are perhaps the most challenging component, but the new GitHub Classroom tool (https://classroom.github.com) allows instructors to host private assignments submitted as code files (.R or .Rmd) or .pdf files, and even build custom autograding tests.

@@

Personal or lab websites are one method of improving the dissemination of research findings and coordination of research efforts, but most EEB researchers have little experience in building or hosting webpages.
GitHub Pages (\url{https://pages.github.com}) is a feature that, when activated for a repository, renders content written in markdown (e.g., \url{https://github.com/SORTEE-Github-Hackathon/main-website}) as an HTML web page with a URL (e.g., \url{https://sortee-github-hackathon.github.io/main-website/}).
This can be useful for sharing reports (e.g.~written with RMarkdown and rendered to HTML) with collaborators, or for more complex projects like personal, project, or lab websites.
GitHub also offers website templates and it is possible to create a website by forking (\protect\hyperlink{definitions}{Box 1}) a repository hosting a GitHub Pages site as a starting point.
Aside from free hosting services, another benefit is that GitHub Pages are autogenerated, meaning that when content is modified in the associated GitHub repository, the website instantly updates {[}\protect\hyperlink{ref-RVetqmsg}{7}{]}.
Though the templates are useful for quickly starting up a new website, users are able to fully customize their Pages websites (for technical details of customizing GitHub Pages site see {[}\protect\hyperlink{ref-MXxgZJ45}{24}{]}).
We emphasize that despite the many benefits of using GitHub pages (free hosting, templates, customization), this avenue for creating a website will often be more time intensive than the out-of-the-box platforms. ``Static websites'' (i.e., websites where every user sees the same content) can be created using Jekyll (\url{https://jekyllrb.com}) and Hugo (\url{https://gohugo.io}), which also include template libraries for websites that can be hosted freely via GitHub pages.
Several R packages such as distill (\url{https://github.com/rstudio/distill}) and blogdown (\url{https://bookdown.org/yihui/blogdown}) have made the interface to these website generators more accessible.
In addition, Quarto (\url{https://quarto.org}) has options for creating websites (\url{https://quarto.org/docs/websites}) as well as the code-free website generator wowchemy (\url{https://wowchemy.com}) which uses Hugo.

@@

GitHub makes it easy to store and share a variety of data files in the cloud.
If a repository is made ``public'' the URL to the repository can be shared freely with others.
However, for a variety of reasons (e.g., privately owned company, ability to make repositories private, accounts can be deleted at will) GitHub is not considered a long-term data or code repository like Zenodo and Figshare {[}\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-tbl:compare}{\textbf{tbl:compare?}}{]}.
Also, unlike long-term repositories, GitHub does not issue Digital Object Identifiers (DOIs) for content uploaded to their servers.
DOIs are persistent and unique alpha-numeric IDs assigned to research products like papers, code, and data.
DOIs allows tracking and citing research products.
For this reason, scientists who share code and data through GitHub are strongly encouraged to also submit GitHub repository content to a long-term data archive {[}\protect\hyperlink{ref-1Du6fzB8g}{25}{]}.
Fortunately, both long-term repositories mentioned above (Zenodo and Figshare) have integrations with GitHub.
After linking one's GitHub repository to Zenodo or Figshare, every time a release (\protect\hyperlink{definitions}{Box 1}) is made, a snapshot of the entire repository is archived with a versioned, citable DOI.
Linking one's GitHub repository with Zenodo, etc. to obtain a DOI helps work become findable, gives proper attribution, and can ensure long-term stability {[}\protect\hyperlink{ref-iIEKCTLU}{26}{]}.

Many researchers believe that their code is not useful because their analysis is context-specific and not explicitly designed for re-use like software.
However, there are many reasons to share data and code beyond re-use.
Even if code lacks detailed annotation, it shows the exact steps taken to conduct an analysis, and therefore provides the most detailed look into how to reproduce a given analysis {[}\protect\hyperlink{ref-uBJwnPbq}{27}{]}.
This is important in light of the reproducibility crisis {[}\protect\hyperlink{ref-1HZdsK5Kn}{28}{]} and will become increasingly important to the collective scientific enterprise as advances in computing power and accessibility unlock the ability to conduct `big data' meta research with data that has already been collected by others.
Failing to include data and code with publications leaves future scientists with fewer open resources from which to understand the world.

@@

GitHub facilitates collaborative coding and asynchronous working because researchers can seamlessly access and contribute to data and code regardless of disparities in schedules or location.
Project work can also be clearly split among team members, giving them the flexibility to contribute when it best fits their schedule.
We acknowledge that there are important differences between the average software developer and EEB researcher using GitHub, and that not all of GitHub's features are useful for research workflows (\protect\hyperlink{tbl:roles}{Table 1}).
However, git's core features such as forking and branching (\protect\hyperlink{definitions}{Box 1}) allow for simultaneous coding on different versions of the same research project, and alternative versions can be discussed and resolved with GitHub.
When EEB projects are contained on the GitHub platform, researchers can stay abreast of progress made by other collaborators without the need for meetings or emails.
Version control {[}\protect\hyperlink{ref-K7nbP1Ty}{29}{]} features also allow users to make progress and changes without worrying about irreparably writing over someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors.
Periodic code review {[}\protect\hyperlink{ref-hm9PaCLD}{30}{]} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

While GitHub is not considered as user-friendly for manuscript development as conventional text processors {[}\protect\hyperlink{ref-4ny1onB0}{8}{]}, it has been substantially improved with recent tools.
Manuscripts can be written on GitHub with Markdown.
Add-ons like HackMD (\url{https://hackmd.io}), can enable real-time collaboration for individual Markdown documents.
Many tools exist for extending Markdown and Pandoc to add formatting features necessary for scientific writing like in-text citations and figure and table cross references.

We wrote this manuscript using Manubot {[}\protect\hyperlink{ref-YuJbg3zO}{31}{]}, a collaborative manuscript platform that uses Markdown for writing and GitHub for storing and tracking changes to a manuscript over time.
Manubot uses a GitHub Actions-based typesetting system to compile individual Markdown files stored in a GitHub repository into a single LaTeX document, which can be displayed in Word, HTML, or PDF formats.
The resulting manuscript can also be compiled using a journal's .tex template to match their formatting requirements.
Since this tool reruns the entire manuscript compilation process with any change to the underlying repository, it can also accomodate continuous integration of code updates into figures and tables with additional GitHub Actions (e.g.~https://github.com/SORTEE-Github-Hackathon/manuscript/tree/main/.github/workflows).
Manubot also allows for straightforward citation management based on URLs or DOIs.
For more examples of Manubot being used for manuscripts, see {[}\protect\hyperlink{ref-YuJbg3zO}{31}{]}.

Since Manubot works on documents in a distributed format, it can be difficult to edit manuscripts for overall flow with only this tool.
We employed hypothes.is (\url{https://hypothes.is}) to write comments on the HTML manuscript document produced by Manubot, which we then addressed by committing changes to the underlying Markdown files via pull requests.
Other authors can reply to the comments to indicate agreement or disagreement, and to note when changes have been made.
Other tools can also be used for version control of scientific manuscripts including R Markdown via the bookdown package (\url{https://bookdown.org/}), jupyter notebooks {[}\protect\hyperlink{ref-i4FOZpjN}{32}{]} and a relatively new tool, Quarto (\url{https://quarto.org}).

>>>>

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/eafc396e1f32d378749c27716bfc058f4802928f/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/eafc396e1f32d378749c27716bfc058f4802928f}{SORTEE-Github-Hackathon/manuscript@eafc396}
on June 8, 2022.

@@

Most scientists, including ecologists and evolutionary biologists, are increasingly dependent on computational tools in their research\textsuperscript{\protect\hyperlink{ref-fJWFe93e}{1}}.
Researchers write and use code as part of their scientific workflow to perform a wide-variety of tasks ranging from data management, data analysis, and study replication, to the application and the development of tools for hypothesis testing.
To maintain code for scientific collaboration requires an efficient and well-documented work-flow\textsuperscript{\protect\hyperlink{ref-1Kqna6l2}{2}}.
To facilitate this process, scientists have been increasingly adopting tools from information and systems technology, such as cloud-based services for documentation and version control (\emph{e.g.}, from the Google Suite, with Docs, Sheets, and Drive; the Microsoft Suite, with Word, Excel, and OneDrive; and GitHub;\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}}).
However, most researchers lack exposure to adequate software development practices and are required to dedicate valuable time and effort to self-teach the use of research-facilitating tools, and thus may be limited in their ability to adhere to adequate standards of scientific code quality and maintenance\textsuperscript{\protect\hyperlink{ref-fJWFe93e}{1},\protect\hyperlink{ref-10SpoByIw}{4},\protect\hyperlink{ref-O6UbstGG}{5}}.
Here, we review and discuss one of the most used web-based platforms for computational version control and collaboration, GitHub, and provide researchers in ecology and evolutionary biology (EEB) with practical workflows aimed at facilitating their scientific code and management process.

With over 73 million registered users, as of 2022, GitHub is the most widely-used web platform for collaborating on computer code\textsuperscript{\protect\hyperlink{ref-nwCtHDCn}{6}}.
GitHub provides a simplified but powerful web interface that allows users to participate in projects, contribute code, report and discuss software bugs, discover existing code and data, and publish new code.
Through version control, users have a detailed, chronological record of the files and directories stored in their repositories (see \protect\hyperlink{definitions}{Box 1};\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}).
This workflow provides a strong and clear advantage over receiving, processing and sending files back-and-forth (e.g.~via email), a process that can easily become challenging and time-consuming in projects extending in time and in the number of collaborators\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8}}.
Through the combination of version control management and the network- and collaboration-based features, GitHub can broadly facilitate openly available source code alongside concomitant collaborative development\textsuperscript{\protect\hyperlink{ref-kEX5dgzK}{9}}.

Git is the version control software system that enables all the collaborative tools available on GitHub.
Although the understanding of basic concepts of Git (such as commit, push, pull, checkout; see \protect\hyperlink{definitions}{Box 1}) is necessary, the GitHub web-based platform and its integrated development environments (such as the GitHub Desktop) allow users to perform most repository and data management operations without opening Git command-line sessions.
Still, GitHub users are encouraged to explore and improve their proficiency in Git when feeling comfortable, so they can feel confident in performing more flexible, complex operations (such as integrating changes from one branch into another branch with \texttt{git\ merge} and \texttt{git\ rebase}). Examples of extensive explanations on Git can be found in journal articles\textsuperscript{\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-PlcxShQU}{10}}, video tutorials\textsuperscript{\protect\hyperlink{ref-NIS0JOW0}{11}}, and books\textsuperscript{\protect\hyperlink{ref-13jOlVcpp}{12}}.

The expansive GitHub user-community and the numerous resources on how to use GitHub have boosted its growing popularity\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-13jOlVcpp}{12}/,\protect\hyperlink{ref-u5aEVE4B}{13}}.
Nevertheless, although multiple articles have encouraged researchers in EEB to adopt GitHub as part of their research process\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-3DKwn1sY}{14}}, its use is still not widespread.
First-time users without formal training in information technology may face steep learning curves because GitHub and its features have been centered on collaboration for software development in information systems\textsuperscript{\protect\hyperlink{ref-139b0pSGc}{15}}.
Moreover, domain-specific perspectives and resources providing tractable examples and practical guidance for researchers in EEB on GitHub are scarce (but see https://ourcodingclub.github.io; https://www.openscapes.org).
A common adoption of GitHub for collaborating on a variety of research tasks can ultimately enable EEB researchers to spend less time on creating novel processes for collaboration and more time on their research\textsuperscript{\protect\hyperlink{ref-ydrk01SR}{16}}.
More importantly, increasing the availability of data and code management standards -- of which GitHub is one increasingly important component -- make research more reproducible and collaborative\textsuperscript{\protect\hyperlink{ref-13QX8XU3J}{17}}.

@@

In EEB, a research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Having multiple copies of these files is important to ensure that research can continue should one copy of a file be accidentally modified or deleted.
Many researchers begin using GitHub as a means to backup their research compendium\textsuperscript{\protect\hyperlink{ref-MwwMapRG}{18}} to a remote server (pull-commit-push, see \protect\hyperlink{definitions}{Box 1}).
The GitHub repository serves as a centralized backup and also allows users to synchronize files and work on any device with internet access.
Because GitHub is built on git, version control features are also available, allowing users to re-visit or even restore previous versions of the repository.
Storing a research compendium as a GitHub repository has additional benefits such as facilitating collaboration, integration with data and code archiving services, and contribution to open science, all discussed in sections below.

Unfortunately, GitHub repositories have file size limits, so they may not be appropriate for synchronizing large data files.
Commits (see \protect\hyperlink{definitions}{Box 1}{]}) greater than 50 MB receive a warning and commits larger than 100 MB are blocked (\textsuperscript{\protect\hyperlink{ref-1Co6ZZjF1}{19}}).
Therefore, it is recommended that large datasets are stored only locally or using other cloud services.

@@

Lab notebooks are the virtual or physical notebooks used to help researchers at any career stage keep track of their research methods or laboratory policies\textsuperscript{\protect\hyperlink{ref-wwHxTOtm}{20}}.
Digital lab notebooks stored in the cloud provide clear benefits given the ease with which documents can be shared with new employees and updated as policy changes or experimental methods are modified\textsuperscript{\protect\hyperlink{ref-10V7x4H4l}{21}}.
Increasingly, researchers are leveraging GitHub's underlying version control system to maintain and share digital lab notebooks\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}}.
At least for aspects of a research project that involve writing code, a GitHub repository is a form of a lab notebook; when changes are made to files in a version controlled repository, the author of those changes makes a commit (\protect\hyperlink{definitions}{Box 1}) accompanied by a message describing the changes and the reason for them. Later, the entire history of commits and their messages are viewable and can be audited similar to a physical lab notebook\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8}}.
GitHub issues (\protect\hyperlink{definitions}{Box 1}) can be used to prioritize lab objectives and goals, as well as track any status updates.
Some EEB labs have even turned their lab notebooks into shareable websites (\url{https://scheuerell-lab.github.io/lab-book}; \url{https://github.com/HuckleyLab/how_we_work}) as a centralized location for all lab resources.

@@

Modern EEB research is highly collaborative, bringing together multidisciplinary teams from various institutions.
On GitHub, collaborators can share feedback, brainstorm ideas, and troubleshoot problems (Figure \ref{fig:github-diagram}).
GitHub has an ``Issues'' feature that allows for discrete tasks and sub-tasks to be identified, assigned to team members, and categorized with custom labels.
The new GitHub ``Discussion'' feature serves as a message board for conversation.
Scripts, commit messages, and pull requests can be linked directly to issues and discussions, providing a clear record of project workflow.
The use of GitHub for all project-related conversation and planning, rather than e-mail or messaging tools, makes it easier to keep track of progress throughout the lifespan of a project.
Unlike emails and messages which can get lost as more new tasks arise, GitHub issues are intentionally closed by repository administrators hiding the issue from view (closed issues remain accessible but not immediately visible).
Fortunately, it is not essential for all team members to have proficiency in git or programming, as users can interact with Issues and Discussions via web browser or e-mail (e-mail responses still get tracked as comments on the focal GitHub issue).
For larger projects with many team members and tens or hundreds of GitHub issues to sort through, project management software like ZenHub, can help prioritize issues and pull requests.
ZenHub's web interface includes a GitHub Issue visualizer where users can organize issues into high priority or backlogged tasks and link issues together when they are related to a shared project goal or milestone.
GitHub is currently beta testing a similar project management feature called GitHub Projects\textsuperscript{\protect\hyperlink{ref-RhBKe0MG}{22}}. GitHub can also be integrated with other project management software such as Slack (https://slack.github.com/) so that teams are notified through a Slack group chat when updates are made to a repository.

@@

Making presentations can be done through most major high-level programming languages such as R, with \texttt{RMarkdown}\textsuperscript{\protect\hyperlink{ref-QqMezOMg}{23}}, Python, with \texttt{python-ppt} (\url{https://python-pptx.readthedocs.io}), and Julia, with \texttt{Remark.jl} (\url{https://juliapackages.com/p/remark}).
Since all these programs work via code bases, they can be version-controlled through git and GitHub.
Once content is made, hosting a course website can be done through GitHub Pages, and there are lots of templates available (e.g., see \url{https://github.com/topics/course-website}).
This way, the course content can be available to enrolled students, as well as a global pool of learners and teachers interested in the course material.
Content can then be delivered via the course website, and/or a GitHub organization with, for example, template repositories for assignments.
Student submissions are perhaps the most challenging component, but the new GitHub Classroom tool (https://classroom.github.com) allows instructors to host private assignments submitted as code files (.R or .Rmd) or .pdf files, and even build custom autograding tests.

@@

Personal or lab websites are one method of improving the dissemination of research findings and coordination of research efforts, but most EEB researchers have little experience in building or hosting webpages.
GitHub Pages (\url{https://pages.github.com}) is a feature that, when activated for a repository, renders content written in markdown (e.g., \url{https://github.com/SORTEE-Github-Hackathon/main-website}) as an HTML web page with a URL (e.g., \url{https://sortee-github-hackathon.github.io/main-website/}).
This can be useful for sharing reports (e.g.~written with RMarkdown and rendered to HTML) with collaborators, or for more complex projects like personal, project, or lab websites.
GitHub also offers website templates and it is possible to create a website by forking (\protect\hyperlink{definitions}{Box 1}) a repository hosting a GitHub Pages site as a starting point.
Aside from free hosting services, another benefit is that GitHub Pages are autogenerated, meaning that when content is modified in the associated GitHub repository, the website instantly updates\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}.
Though the templates are useful for quickly starting up a new website, users are able to fully customize their Pages websites (for technical details of customizing GitHub Pages site see\textsuperscript{\protect\hyperlink{ref-MXxgZJ45}{24}}).
We emphasize that despite the many benefits of using GitHub pages (free hosting, templates, customization), this avenue for creating a website will often be more time intensive than the out-of-the-box platforms. ``Static websites'' (i.e., websites where every user sees the same content) can be created using Jekyll (\url{https://jekyllrb.com}) and Hugo (\url{https://gohugo.io}), which also include template libraries for websites that can be hosted freely via GitHub pages.
Several R packages such as distill (\url{https://github.com/rstudio/distill}) and blogdown (\url{https://bookdown.org/yihui/blogdown}) have made the interface to these website generators more accessible.
In addition, Quarto (\url{https://quarto.org}) has options for creating websites (\url{https://quarto.org/docs/websites}) as well as the code-free website generator wowchemy (\url{https://wowchemy.com}) which uses Hugo.

@@

GitHub makes it easy to store and share a variety of data files in the cloud.
If a repository is made ``public'' the URL to the repository can be shared freely with others.
However, for a variety of reasons (e.g., privately owned company, ability to make repositories private, accounts can be deleted at will) GitHub is not considered a long-term data or code repository like Zenodo and Figshare\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-tbl:compare}{\textbf{tbl:compare?}}}.
Also, unlike long-term repositories, GitHub does not issue Digital Object Identifiers (DOIs) for content uploaded to their servers.
DOIs are persistent and unique alpha-numeric IDs assigned to research products like papers, code, and data.
DOIs allows tracking and citing research products.
For this reason, scientists who share code and data through GitHub are strongly encouraged to also submit GitHub repository content to a long-term data archive\textsuperscript{\protect\hyperlink{ref-1Du6fzB8g}{25}}.
Fortunately, both long-term repositories mentioned above (Zenodo and Figshare) have integrations with GitHub.
After linking one's GitHub repository to Zenodo or Figshare, every time a release (\protect\hyperlink{definitions}{Box 1}) is made, a snapshot of the entire repository is archived with a versioned, citable DOI.
Linking one's GitHub repository with Zenodo, etc. to obtain a DOI helps work become findable, gives proper attribution, and can ensure long-term stability\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{26}}.

Many researchers believe that their code is not useful because their analysis is context-specific and not explicitly designed for re-use like software.
However, there are many reasons to share data and code beyond re-use.
Even if code lacks detailed annotation, it shows the exact steps taken to conduct an analysis, and therefore provides the most detailed look into how to reproduce a given analysis\textsuperscript{\protect\hyperlink{ref-uBJwnPbq}{27}}.
This is important in light of the reproducibility crisis\textsuperscript{\protect\hyperlink{ref-1HZdsK5Kn}{28}} and will become increasingly important to the collective scientific enterprise as advances in computing power and accessibility unlock the ability to conduct `big data' meta research with data that has already been collected by others.
Failing to include data and code with publications leaves future scientists with fewer open resources from which to understand the world.

@@

GitHub facilitates collaborative coding and asynchronous working because researchers can seamlessly access and contribute to data and code regardless of disparities in schedules or location.
Project work can also be clearly split among team members, giving them the flexibility to contribute when it best fits their schedule.
We acknowledge that there are important differences between the average software developer and EEB researcher using GitHub, and that not all of GitHub's features are useful for research workflows (\protect\hyperlink{tbl:roles}{Table 1}).
However, git's core features such as forking and branching (\protect\hyperlink{definitions}{Box 1}) allow for simultaneous coding on different versions of the same research project, and alternative versions can be discussed and resolved with GitHub.
When EEB projects are contained on the GitHub platform, researchers can stay abreast of progress made by other collaborators without the need for meetings or emails.
Version control\textsuperscript{\protect\hyperlink{ref-K7nbP1Ty}{29}} features also allow users to make progress and changes without worrying about irreparably writing over someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors.
Periodic code review\textsuperscript{\protect\hyperlink{ref-hm9PaCLD}{30}} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

While GitHub is not considered as user-friendly for manuscript development as conventional text processors\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8}}, it has been substantially improved with recent tools.
Manuscripts can be written on GitHub with Markdown.
Add-ons like HackMD (\url{https://hackmd.io}), can enable real-time collaboration for individual Markdown documents.
Many tools exist for extending Markdown and Pandoc to add formatting features necessary for scientific writing like in-text citations and figure and table cross references.

We wrote this manuscript using Manubot\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{31}}, a collaborative manuscript platform that uses Markdown for writing and GitHub for storing and tracking changes to a manuscript over time.
Manubot uses a GitHub Actions-based typesetting system to compile individual Markdown files stored in a GitHub repository into a single LaTeX document, which can be displayed in Word, HTML, or PDF formats.
The resulting manuscript can also be compiled using a journal's .tex template to match their formatting requirements.
Since this tool reruns the entire manuscript compilation process with any change to the underlying repository, it can also accomodate continuous integration of code updates into figures and tables with additional GitHub Actions (e.g.~https://github.com/SORTEE-Github-Hackathon/manuscript/tree/main/.github/workflows).
Manubot also allows for straightforward citation management based on URLs or DOIs.
For more examples of Manubot being used for manuscripts, see\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{31}}.

Since Manubot works on documents in a distributed format, it can be difficult to edit manuscripts for overall flow with only this tool.
We employed hypothes.is (\url{https://hypothes.is}) to write comments on the HTML manuscript document produced by Manubot, which we then addressed by committing changes to the underlying Markdown files via pull requests.
Other authors can reply to the comments to indicate agreement or disagreement, and to note when changes have been made.
Other tools can also be used for version control of scientific manuscripts including R Markdown via the bookdown package (\url{https://bookdown.org/}), jupyter notebooks\textsuperscript{\protect\hyperlink{ref-i4FOZpjN}{32}} and a relatively new tool, Quarto (\url{https://quarto.org}).
